SMT-Boosted Security Types for Low-Level MPC

### 2025-05-10 esop25/new-cpj.tex
checkpoint

\section{Declassification Typing}

@@

  \inferrule[Seq]
      {\Gamma_1,\eqs \vdash \cmd_1 : \Gamma_2 \\
        \Gamma_2,\eqs \vdash \cmd_2 : \Gamma_3}
      {\Gamma_1, \eqs \vdash \cmd_1;\cmd_2 : \Gamma_3}
\end{mathpar}

\begin{definition}[Views]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Define:
  $$
  \view(\Gamma,\eqs,X) \defeq \bigcup_{X' \in (\pow(X) - \varnothing)} \ty
     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X'} x : \ty
  $$
\end{definition}

\begin{definition}[Revelation]
  $$
  \begin{array}{ll}
    \revealed(\ty,H) \defeq & \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \mid  \\
      & \qquad \dty{\varnothing}{\eset} \in \ty\ \wedge\\
      & \qquad f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \in \eset\ \wedge \\
      & \qquad \exists 1 \le i \le n . \cid_i \in H} 
  \end{array}
  $$
\end{definition}

\begin{theorem}[Security Typing]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Then $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
  $$
  \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H) = \setit{\idealf}
  $$
\end{theorem}

@@

>>>>

\section{$\minifed$ Declassification Typing}

@@

  \inferrule[Seq]
      {\Gamma_1,\eqs \vdash \prog_1 : \Gamma_2 \\
        \Gamma_2,\eqs \vdash \prog_2 : \Gamma_3}
      {\Gamma_1, \eqs \vdash \prog_1;\prog_2 : \Gamma_3}
\end{mathpar}
We note that typing tracks uniformity of marginal distributions of independent messages,
but not their joint distributions. In order to capture this, we define a notion of
dependency closures in views:  
\begin{definition}[Views]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Define:
  $$
  \view(\Gamma,\eqs,X) \defeq \bigcup_{X' \in (\pow(X) - \varnothing)} \ty
     \qquad \text{where}\ \Gamma,\eqs \vdash \Sigma_{x \in X'} x : \ty
  $$
\end{definition}
The goal of the analysis is to show that views contain only noise and the
result of the ideal functionality. To this end we define a notion of revelation, where
functions of honest secrets are available in the clear in adversarial
views:
\begin{definition}[Revelation]
  $$
  \begin{array}{ll}
    \revealed(\ty,H) \defeq & \setit{f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \mid  \\
      & \qquad \dty{\varnothing}{\eset} \in \ty\ \wedge\\
      & \qquad f(\sx{w_1}{\cid_1},\ldots,\sx{w_n}{\cid_n}) \in \eset\ \wedge \\
      & \qquad \exists 1 \le i \le n . \cid_i \in H} 
  \end{array}
  $$
\end{definition}
Now we can state our security typing correctness result, which says that the only
information revealed in adversarial views (besides noise) is the ideal functionality. 
\begin{theorem}[Security Typing]
  Given $\prog, \eqs, \Gamma$ with $\toeq{\prog} \models \eqs$ and $\varnothing, \eqs \vdash \prog : \Gamma$.
  Then $\prog$ satisfies {NIMO} for ideal functionality $\idealf$ if for all $H,C$, we have:
  $$
  \revealed(\view(\Gamma,\eqs,\vars(\prog)_C),H) = \setit{\idealf}
  $$
\end{theorem}

@@

The type system presented above is in a logical form and is not syntax-directed.
Also, the definition of $\view$ involves a powerset construction with exponential
complexity in the size of the program.

As a step towards an algorithm, we can define specific strategies and analysis
based on our type system that both reduces overall complexity and leverages SMT
solving. This will form the foundation of a decidable and scalable type analysis
for $\metaprot$.  We define \emph{vectors} of messages $\nu$ that participate in strategies,
e.g., the collective shares of secret values.
\begin{definition}
  A \emph{message vector} $\nu$ is either a vector of distinct
  broadcast messages $\cvec{\px{w_1},\ldots,\px{w_n}}$, or of unicast
  messages $\cvec{\mx{w_1}{\cid_1},\ldots,\mx{w_n}{\cid_n}}$ where
  each $\cid_i$ for $i \in \setit{\cid_1,\ldots,\cid_n}$ is distinct.  
\end{definition}

\emph{Strategy typings} and environments are then defined via the following syntax.
$$
\begin{array}{rcl@{\hspace{4mm}}r}
\sty &::=& \sharet{\phi} \mid \declasst{\phi} \mid \secopent & \gdesc{Strategy Types} \\[1mm]
\Delta &::=& \varnothing \mid \Delta; \nu : \sty & \gdesc{Strategy Type Environments}
\end{array}
$$
